# Introduction to Data Science HW 4
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Sana Khan, HW 4 Due 1/8
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
#library(jsonlite)
#dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
#readlines <- jsonlite::fromJSON(dataset)
#df <- readlines$objects$person
#This code is installing and activating the jsonlite package. It's then assigning the url to the dataset variable. The next line is taking the variable and using the fromJSON function from the jsonlite package to import the data into R. The last line is storing it in a new variable by pulling just the person data from the objects.
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}
str(df)
library(tidyverse)
head(df)
glimpse(df)
```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}
#This data set is about US Senators. There are 1000 rows and each row represents a Senator. There are 17 columns which represent characteristics about the senators. 
```

C. What does running this line of code do? Explain in a comment:


```{r}
vals <- substr(df$birthday,1,4)
#It finds the birthdays for the first four senators. The code is picking rows 1-4. 
```

D. Create a new attribute 'age' - how old the person is 
**Hint:** You may need to convert it to numeric first.


```{r}
df$age = as.numeric(difftime(Sys.Date(),df$birthday, units = "weeks"))/52


```

E. Create a function that reads in the role json dataset, and adds the age attribute to the dataframe, and returns that dataframe


```{r}
my_func <- function(url_link){
  dataset <- url(url_link)
readlines <- jsonlite::fromJSON(dataset)
df <- readlines$objects$person
df$age = as.numeric(difftime(Sys.Date(),df$birthday, units = "weeks"))/52
return(df)
}

```

F. Use (call, invoke) the function, and store the results in df


```{r}
df <-my_func("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")

```

## Part 2: Investigate the resulting dataframe 'df'

A.	How many senators are women? 


```{r}
table(df$gender)
#24 senators are women
```

B.	How many senators have a YouTube account?  


```{r}
df %>%
  summarise(sum(!is.na(youtubeid)))
```

C.	 How many women senators have a YouTube account?


```{r}
df  %>%
  group_by(df$gender) %>%
   summarise(sum(!is.na(youtubeid))) 
        #16 Female senators have a youtube account
```

D.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
youtubeWomen <- filter(df, gender == 'female',!is.na(youtubeid) )

```

E.	Make a histogram of the **age** of senators in **youtubeWomen**, and then another for the senetors in **df**. Add a comment describing the shape of the distributions.


```{r}
hist(youtubeWomen$age)
hist(df$age)
#The distribution of the age of all the senators have a relatively normal distribution, where as the distribution of the age of female senators skews to the left. This can be attributed to the population size, there are more samples in the df dataframe than in the women senators dataframe. 
```
