# Intro to Data Science - HW 6
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Sana Khan 
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.

```

Last assignment we explored **data visualization** in R using the **ggplot2** package. This homework continues to use ggplot, but this time, with maps.  In addition, we will merge datasets using the built-in **merge( )** function, which provides a similar capability to a **JOIN in SQL** (don't worry if you do not know SQL). Many analytical strategies require joining data from different sources based on a **“key”** – a field that two datasets have in common. 

## Step 1: Load the population data

A.	Read the following JSON file, https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json and store it in a variable called **pop**.

Examine the resulting pop dataframe and add comments explaining what each column contains.



```{r, warning=FALSE, message=FALSE}

library(tidyverse)
library(jsonlite)
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json")
pop <- jsonlite::fromJSON(dataset)
pop

#This dataframe shows data on cities that experienced population growth from 2000 - 2013. The dataframe contains the city, state, lat and long as well as population and rank
```

B.	Calculate the **average population** in the dataframe. Why is using mean() directly not working? Find a way to correct the data type of this variable so you can calculate the average (and then calculate the average)

Hint: use **str(pop)** or **glimpse(pop)** to help understand the dataframe


```{r}
pop$population <-as.numeric(pop$population)
mean(pop$population)
 #The mean is  131132.4

```

C.	What is the population of the smallest city in the dataframe? Which state is it in?


```{r}
pop[which.min(pop$population),]

#Population is 36877 and is for Panama City

```

## Step 2: Merge the population data with the state name data

D)	Read in the state name .csv file from the URL below into a dataframe named **abbr** (for "abbreviation") – make sure to use the read_csv() function from the tidyverse package: <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv



```{r}
abbr<- read_csv(url("https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv"))
abbr
```

E)	To successfully merge the dataframe **pop** with the **abbr** dataframe, we need to identify a **column they have in common** which will serve as the **“key”** to merge on. One column both dataframes have is the **state column**. The only problem is the slight column name discrepancy – in **pop**, the column is called **“state”** and in **abbr** – **“State.”** These names need to be reconciled for the merge() function to work. Find a way to rename **abbr’s “State”** to **match** the **state column in pop**.    


```{r}
abbr <- abbr %>%
  rename(state=State)

abbr
```

F)	Merge the two dataframes (using the **‘state’ column** from both dataframes), storing the resulting dataframe in **dfNew**.


```{r}
dfNew <- merge(abbr, pop, by= 'state')
dfNew 
```

G)	Review the structure of **dfNew** and explain the columns (aka attributes) in that dataframe.


```{r}
str(dfNew)
#Merging the two data frames added the State and Abbreviation colums to the data set that had the city, population and longitude and latitude data
```

## Step 3: Visualize the data

H)	Plot points (on top of a map of the US) for **each city**. Have the **color** represent the **population**.


```{r, warning=FALSE, message=FALSE}

library(ggmap)


bb <- c(left = min(dfNew$long),
bottom = min(dfNew$lat),
right = max(dfNew$long),
top = max(dfNew$lat))

map = get_stamenmap(bb, zoom = 4)

ggmap(map)

ggmap(map) +
geom_point(data= dfNew,
aes(x=longitude,y=latitude, color=population))


```

I)	Add a block comment that criticizes the resulting map. It’s not very good.

```{r}
#Because the plots are reprenting the color and the population it makes it difficult to truly understand what the map is trying to show
```

## Step 4: Group by State


J)	Use group_by and summarise to make a dataframe of state-by-state population. Store the result in **dfSimple**.


```{r}
dfSimple <- dfNew %>%
  group_by(state)%>%
  summarise(total_population = sum(population))
dfSimple
```

K)	Name the most and least populous states in **dfSimple** and show the code you used to determine them.


```{r}
summary(dfSimple)

dfSimple[which.min(dfSimple$total_population),]
#vermont
dfSimple[which.max(dfSimple$total_population),]
#california
```

## Step 5: Create a map of the U.S.,  with the color of the state representing the state population

L) Make sure to expand the limits correctly and that you have used **coord_map** appropriately.



```{r}
library(ggmap)
us <-map_data("state")
head(us)

dfSimple$state <- tolower(dfSimple$state)
head(dfSimple)

combined_data <- merge(us, dfSimple, by.x="region", by.y= 'state')
head(combined_data)


ggmap(map) +
geom_polygon(data=combined_data, aes(x= long,y=lat, group=group, fill=total_population),color="black", alpha=0.8,) +
  expand_limits(x=combined_data$long, y=combined_data$lat)



```
